This work explored a solution to the problem faced by bioengineers to determine those cellular mechanisms to target for modification when improving vascularization for synthetic tissue engineering.  The method integrated a hybrid particle-based model of vascularization as a fitness function with a genetic algorithm applied to optimize the search space of model parameter values. The method is generic in that models of different biological systems could be substituted for vascularization and the fitness function likewise changed to reflect alternative objectives, such as minimizing vascular function to combat tumor growth in cancer \cite{Mahoney2010MultiObjective}.

In this initial study, the search space was reduced to only consider parameters that influence the chemotactic mechanism of the vascular cells. Previous work with the model identified the importance of adhesion and tight junction formation in robust network formation. However, increasing the number of parameters to include these cellular mechanisms would significantly increase the search space, resulting in many more fitness function calls. Given the modeling framework employed here \cite{Lardon2011IDynoMiCS}, this is currently impractical because each fitness evaluation takes at least four hours of CPU time running on a 3.6~GHz machine. Recently two fast large-scale simulation systems have been developed by Ghaffarizadeh et al. \cite{ghaffarizadeh2015agent} and \textsl{Biocellion} \cite{delindavis:biocellion}. Both these systems implement an individual-based approach similar to \textsl{cDynoMiCs} employed here. \textsl{Biocellion} is implemented as a distributed architecture executable on the Cloud \cite{Hashem2015Rise} and is capable of simulating complex 3D models of billions of cells in a matter of a few hours. \textsl{Biocellion} has the potential to simulate vascularization quickly enabling the optimization of more complex bioengineered multicellular systems.

The principle challenge in bringing this work to practicality is to operationalize the link between model parameters and actual genetic sequences. If this can be achieved, then the output of the optimization process could be directly mapped to engineering targets in the genome. In traditional bioengineering where products are produced in biofactories \cite{delindavis:Sharma2001Production}, \cite{delindavis:vanDijl2013Bacillus}, targets are identified through analysis of detailed metabolic models of microbial cells, such as \cite{Karp2015Pathway}. The functioning of mammalian cells is less understood, particularly mechanisms such as chemotaxis that are unrelated to metabolic processes, which can be modeled using flux balance analysis.

The key to operationalizing bioengineering of multicellular systems is to utilize multiscale models of the biological system under study. Multiscale models link the mechanisms acting at different spatial and temporal scales together into an integrated system where changes in expression and regulation of genes are manifested in large-scale multicellular outcomes \cite{Martins2010Multiscale}, \cite{JosephWalpole2013Multiscale} \cite{Yu2016Multiclass}. If such detailed models were employed, the high-level abstract parameters of this work would be replaced by specific genetic mechanisms resulting in solutions being mapped directly to genetic targets.


