The self organizing vascular model has seven parameters with unknown values (Table~\ref{unknownparameters}).  The best fit for these parameters was found using a genetic algorithm from the Java Genetic Algorithms Package \cite{cgl:meffert2011genetic}.  The seven parameters are used as genes in individual chromosomes.  The initial population had a size of three and was constructed by random selection of parameter values from the ranges shown in Table~\ref{unknownparameters}. As the population evolves during runs of the genetic algorithm, each individual is evaluated using the parameter values encoded in the genes used to run the vascular model (see Section~\ref{vascularModel} for more details). The resulting vascular network is then evaluated as described in Section~\ref{fitnessFunction}, and the fitness is assigned based on the metabolic production. Next, genetic operators were run on the population followed by a re-evaluation of the population that now contained new individuals. Finally, selection was applied to obtain the new population for the next round of evolution.
Crossover and mutation operators were included in the genetic algorithm, with the crossover operator applied to 35\% of the population per generation and the mutation operator was applied to each gene with a probability of 1/12. An elitist ranking selector cloned the top \%90 of the population, with the remaining 10\% of the population filled with the same individuals in the order of fitness. The results presented in this paper were obtained using 1000 steps of evolution.


\begin{table}[ht]
\caption{Parameters searched by the GA} % title of Table
\centering
\begin{footnotesize}
\begin{tabular}{l l l l}
\hline
Parameter        & Low & High & Description\\ \hline \hline
$k$              & $10^{-10}$ & $10^{-2}$ & Monod-kinetic coefficient \\
$\mu_{short}$     & $1$ & $7$             & Secretion rate of $C_{short}$ \\
$\mu_{long}$      & $1$ & $7$             & Secretion rate of $C_{long}$ \\
$\beta_{short}$   & $10^{-5}$ & $10^{-1}$  & Decay rate of $C_{short}$ \\
$\beta_{long}$    & $10^{-10}$ & $10^{-3}$ & Decay rate of $C_{long}$ \\
$\lambda_{short}$ & $1$ & $6$             & Chemoattractant response to $C_{short}$ \\
$\lambda_{long}$  & $1$ & $6$             & Chemoattractant response to $C_{long}$ \\
\hline
\end{tabular}
\end{footnotesize}
\label{unknownparameters}
\end{table}
%% Range Values from VNoptimization.java, VNFitnessFunction.java
%%
%% gene[0] = new DoubleGene(gaConf, 1, 6);                  lambda_short
%% gene[1] = new DoubleGene(gaConf, 1, 6);                  lambda_long
%% gene[2] = new DoubleGene(gaConf, 1, 7 );                 mu_short
%% gene[3] = new DoubleGene(gaConf, 1, 7 );                 mu_long
%% gene[4] = new DoubleGene(gaConf, 0.00001, 0.1 );         beta_short
%% gene[5] = new DoubleGene(gaConf, 0.0000000001, 0.001 );  beta_long
%% gene[6] = new DoubleGene(gaConf, 0.0000000001, 0.01);    k
%%

